\normalfont{A bird's-eye view on the habitability of exoplanets via statistical learning techniques}

### 2020-11-03 Manuscript/Manuscript.tex
PUB

\subsection{Decision tree and random forest}
The first approach that was considered is the classification decision tree \footnote{In this work, for all the approaches undertaken, the training ensemble consist in a set of 350 exoplanets while the test one is composed by 150 items.}: as it can be seen from Fig \ref{DecisionTree}, where the decision tree is given accompanied by a plot that provides the variables importance,  the main discriminant parameters, as expected, are the planet temperature equilibrium $P\_T\_E$ and the stellar temperature $S\_T$: this features discriminates, as one expects, the largest part of the habitable planet from the non-habitable. The performance of this algorithm on this dataset, evaluated with a 10-fold cross-validation \footnote{For this procedure, applied to all algoritms here considered the test set and the validation sets where reshuffled 10 times by modifying the numerical seed by which the exoplanets are assigned to the training or to the test set. The proportion 150/500 for test and training sets was kept constant}, is reported in Fig \ref{Decision_tree_performances} and in the table \ref{perf_table}. Next, in order to improve the performance of this classification, the random forest was considered. As shown in Fig. \ref{Random_forest_tree_perfs} the main parameters for this algorithm were previously checked. Its results are reported in Fig \ref{Random_forest_tree_results}: here the most important variable is still $P\_T\_E$ followed by the $S\_T$.  Moreover  as one can note from Fig \ref{Random_forest_tree_results} the performances are slightly improved with respect to the ones reported in Fig \ref{Decision_tree_performances}.  Such improvement is captured by the accuracy and $\phi$ factors reported in table \ref{perf_table}. Finally it is worth noting that, although  different features were also considered and the output of the response function has a larger set,  a similar hierarchy  among the variable importance for the random forest algorithm was obtained by Saha et al. \cite{saha2018machine}: in their paper the highest variable importance, via random forest algorithm,  is assigned to the mean planet temperature at surface followed by the stellar magnitude with respect to the planet (between these two there are other features strictly connected to the planet surface temperature). The last quantity is strictly connected to the stellar temperature and to the stellar spectral type. Moreover also in their study the mean distance between the planet and host star plays a minor role.

>>>>

\subsection{Decision tree and random forest}
The first approach that was considered is the classification decision tree \footnote{In this work, for all the approaches undertaken, the training ensemble consist in a set of 350 exoplanets while the test one is composed by 150 items.}: as it can be seen from Fig \ref{DecisionTree}, where the decision tree is given accompanied by a plot that provides the variables importance,  the main discriminant parameters, as expected, are the planet temperature equilibrium $P\_T\_E$ and the stellar temperature $S\_T$: this features discriminates, as one expects, the largest part of the habitable planet from the non-habitable. The performance of this algorithm on this dataset, evaluated with a 10-fold cross-validation \footnote{For this procedure, applied to all algorithms here considered, the test set and the training sets where reshuffled 10 times by modifying the numerical seed by which the exoplanets are assigned to the training or to the test set. The model was than trained again with the reshuffled set. The confusion matrices reported here are the sum of the 10 forecast on the test set obtained.}, is reported in Fig \ref{Decision_tree_performances} and in the table \ref{perf_table}. Next, in order to improve the performance of this classification, the random forest was considered. As shown in Fig. \ref{Random_forest_tree_perfs} the main parameters for this algorithm were previously checked. Its results are reported in Fig \ref{Random_forest_tree_results}: here the most important variable is still $P\_T\_E$ followed by the $S\_T$.  Moreover  as one can note from Fig \ref{Random_forest_tree_results} the performances are slightly improved with respect to the ones reported in Fig \ref{Decision_tree_performances}.  Such improvement is captured by the accuracy and $\phi$ factors reported in table \ref{perf_table}. Finally it is worth noting that, although  different features were also considered and the output of the response function has a larger set,  a similar hierarchy  among the variable importance for the random forest algorithm was obtained by Saha et al. \cite{saha2018machine}: in their paper the highest variable importance, via random forest algorithm,  is assigned to the mean planet temperature at surface followed by the stellar magnitude with respect to the planet (between these two there are other features strictly connected to the planet surface temperature). The last quantity is strictly connected to the stellar temperature and to the stellar spectral type. Moreover also in their study the mean distance between the planet and host star plays a minor role.
